Gusto: a toolkit for compatible finite element dynamical cores

### 2025-08-05 source/paper.tex
finish transport section

>>>>

The starting point for this discretisation is to replace the
$\MM{u}\cdot\nabla\MM{u}$ term with $(\nabla\times\MM{u})\times\MM{u}
+ \frac{1}{2}\nabla(\MM{u}\cdot\MM{u})$. Now, multiplying by a test
function and integrating by parts over an element $e$ with boundary
$\partial e$ gives
\begin{equation}
  \int_e \MM{w}\cdot(\nabla\times\MM{u})\times\MM{u}
  + \frac{1}{2}\MM{w}\cdot\nabla(\MM{u}\cdot\MM{u}) \diff x = \int_e\nabla\times(\MM{u}\times\MM{w})\cdot\MM{u}\diff x - \int_{\partial e}\MM{n}\times(\MM{u}\times\MM{w})\cdot\MM{u} \diff S - \frac{1}{2}\int_e\nabla\cdot\MM{w}\MM{u}\cdot\MM{u}\diff x, \quad \forall \MM{w} \in \mathbb{V}^{HDiv}.
\end{equation}
Sunming over the domain and taking the upwind value of $\MM{u}$,
denoted by $\tilde{\MM{u}}$, on the element boundaries gives
\begin{equation}
  \int_\Omega \MM{w}\cdot(\nabla\times\MM{u})\times\MM{u}
  + \frac{1}{2}\MM{w}\cdot\nabla(\MM{u}\cdot\MM{u}) \diff x = \int_\Omega\nabla\times(\MM{u}\times\MM{w})\cdot\MM{u}\diff x - \int_{\Gamma}\jump{\MM{n}\times(\MM{u}\times\MM{w})}\cdot\tilde{\MM{u}} \diff S - \frac{1}{2}\int_\Omega\nabla\cdot\MM{w}\MM{u}\cdot\MM{u}\diff x, \quad \forall \MM{w} \in \mathbb{V}^{HDiv},
\end{equation}
where $\tilde{\MM{u}}$ denotes the upwind value of $\MM{u}$ on the facet.
